---
title: "Homework 2 Report"
output: html_document
---


## Name

Caroline Zhuang

### Setup

Read in the data with `read_csv()` and store the data as an R object named `dataset`. Check the data to make sure all of the expected observations and variables are there.


```{r}
## Load the data and any necessary packages here.
library(ggplot2)
library(dplyr)
library(tidyr)
dataset <- read.csv('maacs.csv.gz')
dataset
```

## Part 1

We will first consider the relationship between FEV1 and age. In general, it is expected that as children get older (and hence, larger in size), their FEV1 values should get higher. 

Consider the statement "FEV1 values in children are higher in older children relative to younger children". 

Write a function in R that takes the `dataset` object as an argument and returns `TRUE` if the statement above is true for the dataset and `FALSE` otherwise. 

NOTE: In order to write this function, you will need to translate the statement above into something that can be checked with the data. There are many ways in which you can do that translation correctly and you only need to pick one way here. 

NOTE: For this part, do not use any plots.

```{r}
## Write your function here
are_fev1_higher_older <- function(dataset) {
  cor_val <- cor(dataset$age, dataset$fev1, use = "complete.obs")
  return(cor_val > 0)
}
```

```{r}
answer_q1 <- are_fev1_higher_older(dataset)
answer_q1
```


## Part 2

Fit a linear regression model with FEV1 as the outcome and age as a predictor.

How much does FEV1 change for a 1-year increase in the child's age? 

```{r}
## Add your code here
x <- dataset$age
y <- dataset$fev1
mod <- lm(y ~ x)
cf <- coef(mod)
cf[1]
cf[2]
```


Write your data analysis statement interpreting the regression model here:

When the ages of the subjects happen to differ by 1 year, there is an increase 
of FEV1 by around 0.17 liters. When the age of the subjects is 0, the FEV1 value
is -0.0012 liters, which both are impossible values, so the intercept does not 
provide any context in this scenario.

## Part 3

Develop **three** supporting premises derived from the data that support the statement you wrote in Part 2. These can be plots, other summary statistics, or model results. 

NOTE: 

* At least one supporting premise should use a plot.

* Do not use the code you write in Part 1 as a supporting premise

```{r}
## Add your code here
# defining the subgroups 
median_age <- median(dataset$age)
young_children <- subset(dataset, age <= median_age)
older_children <- subset(dataset, age > median_age)

# statement 1 
mean_young <- mean(young_children$fev1, na.rm = TRUE)
mean_old <- mean(older_children$fev1, na.rm = TRUE)

# statement 2
dataset$group <- ifelse(dataset$age <= median_age, "Young", "Old")
dist_plot <- ggplot(dataset, aes(group, fev1)) +
  geom_boxplot() 
dist_plot

# statement 3 
hist_plot <- ggplot(dataset, aes(fev1)) +
  geom_histogram(bins = 20) +
  facet_wrap(~group, ncol = 1)
hist_plot
```


Write the three supporting premise statements here:

1. The mean FEV1 among older children is higher than the mean FEV1 value in 
younger children.

2. The median FEV1 among older children is higher than the median FEV1 value in
younger children, as shown by the boxplot. 

3. The distribution of FEV1 values shifts higher for older children compared to
younger children, as shown in the histogram.


## Part 4

For each of the supporting premises above, write a function that takes the `dataset` object as an argument and returns `TRUE` if the supporting premise statement above is true for the dataset and `FALSE` otherwise. 

For statements involving plots, instead of returning `TRUE` or `FALSE`, you function should do two things:

1. Produce the plot that is used in the statement

2. Produce a hypothetical version of the plot in the event that the statement is true. This can be done using simulated data or by simply hand drawing a plot.


```{r}
## Function for supporting premise statement 1
mean_higher <- function(dataset){
  # defining the subgroups 
  median_age <- median(dataset$age)
  young_children <- subset(dataset, age <= median_age)
  older_children <- subset(dataset, age > median_age)
  
  # statement 1 
  mean_young <- mean(young_children$fev1, na.rm = TRUE)
  mean_old <- mean(older_children$fev1, na.rm = TRUE)
  
  return(mean_old > mean_young)
}

mean_higher(dataset)
```


```{r}
## Function for supporting premise statement 2
boxplot_groups <- function(dataset){
  # defining groups 
  median_age <- median(dataset$age)
  dataset$group <- ifelse(dataset$age <= median_age, "Young", "Old")
  
  # true data
  true_dist_plot <- ggplot(dataset, aes(group, fev1)) +
    geom_boxplot() +
    labs(title = "True: Median FEV1 By Group",
         x = "Group", y = "FEV1")
  
  # simulated data
  set.seed(123)
  n <- 50  # number per group
  young_sim <- rnorm(n, mean = 2.0, sd = 0.4)   # lower center
  old_sim   <- rnorm(n, mean = 3.0, sd = 0.4)   # higher center
  
  sim_data <- tibble(
    group = rep(c("Young", "Old"), each = n),
    fev1  = c(young_sim, old_sim)
  )
  
  simulated_plot <- ggplot(sim_data, aes(x = group, y = fev1)) +
    geom_boxplot() +
    labs(title = "Hypothetical: Median Higher in Older",
         x = "Group", y = "FEV1")
  
  return(list(true_dist_plot, simulated_plot))
}

box_plots <- boxplot_groups(dataset)
box_plots
```


```{r}
## Function for supporting premise statement 3
hist_groups <- function(dataset){
  # defining groups 
  median_age <- median(dataset$age)
  dataset$group <- ifelse(dataset$age <= median_age, "Young", "Old")
  
  # true data
  true_hist_plot <- ggplot(dataset, aes(fev1)) +
    geom_histogram(bins = 20) +
    facet_wrap(~group, ncol = 1) +
    labs(title = "True Distribution of FEV1 by Group", 
         x = "Group", y = "FEV1")
  
  # simulated data
  set.seed(123)
  n <- 50  # number per group
  young_sim <- rnorm(n, mean = 2.0, sd = 0.4)   # lower center
  old_sim   <- rnorm(n, mean = 3.0, sd = 0.4)   # higher center
  
  sim_data <- tibble(
    group = rep(c("Young", "Old"), each = n),
    fev1  = c(young_sim, old_sim)
  )
  
  simulated_plot <- ggplot(sim_data, aes(fev1)) +
    geom_histogram(bins = 20) +
    facet_wrap(~group, ncol = 1) +
    labs(title = "Hypothetical: Shifted Distribution in Older",
         x = "Group", y = "FEV1")
    
  return(list(true_hist_plot, simulated_plot))
}

hist_plots <- hist_groups(dataset)
hist_plots
```

Execute each of your function and show that the produce the expected output.


## Part 5

Describe one alternative to the primary statement "FEV1 values in children are higher in older children relative to younger children". 

Create a fault tree for the alternative outcome describing how the alternative outcome could be realized in the data even if the primary statement were true.

Your fault tree should be created as a separate image and does not need to be created in R. Upload the image of the fault tree to Canvas.


